---
title: "Taxonomic barplots with >1% abundance"
author: "Nat Pombubpa"
date: "Mar 10, 2020"
output: html_document
---
#Months

```{r warning=FALSE, message=FALSE}
library(ape)
library(vegan)
library(plyr)
library(dplyr)
library(scales)
library(grid)
library(reshape2)
library(phyloseq)
library(magrittr)
library(ggplot2)
library(ggpubr)
library(data.table)
library(tidyr)
library(tidyverse)
library(multcompView)
library(VennDiagram)
library(car)
#library(Hmisc)
```

###STEP1: Import otu table for analysis using Phyloseq and drop taxonomy column
```{r warning=FALSE}
otus.raw = read.table("emp_streptomyces_w_taxonomy.txt",
                   header=T,sep="\t",row.names = 1)
otus.raw.notax = otus.raw[1:(length(otus.raw)-1)]
```

```{r}
otumat <- as(as.matrix(otus.raw.notax), "matrix")
OTU = otu_table(otumat, taxa_are_rows = TRUE)
```

###STEP2: Import taxonomy table for analysis using Phyloseq
select taxnomy column (last column)
```{r warning=FALSE}
#taxmat.raw = read.table("emp_streptomyces_w_taxonomy.txt",
#                   header=T,sep="\t",row.names = 1)
taxmat.raw.noOTU = otus.raw[,ncol(otus.raw), drop = FALSE]
```

Brute-force remove D_0__, .....
```{r}
taxmat.raw.noOTU = as.data.frame(taxmat.raw.noOTU)
taxmat.raw.noOTU$taxonomy = gsub("D_0__","",as.character(taxmat.raw.noOTU$taxonomy))
taxmat.raw.noOTU$taxonomy = gsub("D_1__","",as.character(taxmat.raw.noOTU$taxonomy))
taxmat.raw.noOTU$taxonomy = gsub("D_2__","",as.character(taxmat.raw.noOTU$taxonomy))
taxmat.raw.noOTU$taxonomy = gsub("D_3__","",as.character(taxmat.raw.noOTU$taxonomy))
taxmat.raw.noOTU$taxonomy = gsub("D_4__","",as.character(taxmat.raw.noOTU$taxonomy))
taxmat.raw.noOTU$taxonomy = gsub("D_5__","",as.character(taxmat.raw.noOTU$taxonomy))
taxmat.raw.noOTU$taxonomy = gsub("D_6__","",as.character(taxmat.raw.noOTU$taxonomy))
```

```{r}
taxmat.raw.noOTU
```


Separate taxon into different levels
```{r}
tax_filtered = separate(taxmat.raw.noOTU, taxonomy, c("Kingdom","Phylum","Class","Order", "Family", "Genus","Species"), sep= ";", remove=TRUE)
```

```{r warning=FALSE}
taxmat <- as(as.matrix(tax_filtered),"matrix")
TAX = tax_table(taxmat)
```

```{r}
head(TAX)
```


###STEP3: Import mapping file for analysis using Phyloseq

1.Check mapping file before import to R, R will automatically change sample name that starts with number or contain “-” in sample name. If you get error in this step, you should check sample name first. 

2.First column of first row should not start with #, R will not read the first row that starts with #

3. You can choose which samples to include in analysis by indicating specific group in the column

```{r warning=FALSE}
meta = read.csv("emp_qiime_mapping_release1_20170912_cp.tsv",
                  header=TRUE,row.names=1,
                  sep="\t",stringsAsFactors=FALSE)
#meta <- meta[which(meta$Layer %in% c("Surface")),]
```

Check if your metadata file has been import successfully and correctly, the output will show a table of your metadata file (mapping file). *If you do not have header, you might start your first row with # (remove # and reload your mapping file).

```{r}
head(meta)
```

Construct sample_data-class using imported metadata
```{r warning=FALSE}
sampleData <- sample_data(meta)
```

###STEP4: Construct Phyloseq object
To construct phyloseq object, otu table, taxonomy table, and sampleData are required. Phylogenetic tree can be included, but it is not necessary for constructing phyloseq object.
Construct Phyloseq object called "Physeq"

```{r warning=FALSE}
physeq = phyloseq(OTU,TAX,sampleData)
```

###STEP8: Check phyloseq object
This should indicate that your physeq is a "phyloseq-class experiment-level object"

```{r}
physeq
```



remove any OTUs that present only one time (singletons)

```{r}
physeq.prune = prune_taxa(taxa_sums(physeq) > 1, physeq)
```

```{r}
physeq.prune
```

remove any samples that has only one read
```{r}
physeq.prune.sample = prune_samples(sample_sums(physeq.prune) >= 5, physeq.prune)
physeq.prune.sample
```

Remove negative samples
```{r}
physeq.prune.sample.noneg = subset_samples(physeq.prune.sample, empo_2!="Negative")
physeq.prune.sample.noneg
```

Remove " uncultured bacterium"
```{r}
physeq.prune.sample.noneg.notaxa = subset_taxa(physeq.prune.sample.noneg, Species != " uncultured bacterium")
physeq.prune.sample.noneg.notaxa
```

Remove " bacterium NR3"
```{r}
physeq.prune.sample.noneg.notaxa1 = subset_taxa(physeq.prune.sample.noneg.notaxa, Species != " bacterium NR3")
physeq.prune.sample.noneg.notaxa1
```

Remove " Actinomycetales bacterium HPB43"
```{r}
physeq.prune.sample.noneg.notaxa2 = subset_taxa(physeq.prune.sample.noneg.notaxa1, Species != " Actinomycetales bacterium HPB43")
physeq.prune.sample.noneg.notaxa2
```

Remove " Actinomycetales bacterium HO2017"
```{r}
physeq.prune.sample.noneg.notaxa3 = subset_taxa(physeq.prune.sample.noneg.notaxa2, Species != " Actinomycetales bacterium HO2017")
physeq.prune.sample.noneg.notaxa3
```

Remove " uncultured actinobacterium"
```{r}
physeq.prune.sample.noneg.notaxa4 = subset_taxa(physeq.prune.sample.noneg.notaxa3, Species != " uncultured actinobacterium")
physeq.prune.sample.noneg.notaxa4
```

```{r}
physeq.prune.sample.noneg.notaxa4
```

remove any samples that has less than 100 reada
```{r}
physeq.beta.100 = prune_samples(sample_sums(physeq.prune.sample.noneg.notaxa4) >= 100, physeq.prune.sample.noneg.notaxa4)
physeq.beta.100
```

CLR transform
```{r}
physeq_clr = microbiome::transform(physeq.beta.100, "clr")  
```

Non CLR
```{r}
phyloseq::otu_table(physeq.beta.100)[1:5, 1:5]
```

CLR 
```{r}
phyloseq::otu_table(physeq_clr)[1:5, 1:5]
```

#Figure 5: Beta Div Overall

STOP here, results from HPCC showed that there might be some NA in otu table. need to find a way to change NA to 0

```{r}
physeq_clr_ord <- ordinate(physeq.beta.100, "PCoA", "bray")
```

```{r}
sample_data(physeq_clr)$Month = factor(sample_data(physeq_clr)$Month, levels = c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December"))
```

```{r}
colnames(meta)
```

```{r}
ps.dist = phyloseq::distance(physeq.beta.100, "bray")
```

```{r}
adonis(ps.dist ~empo_2+country+envo_biome_2, as(sample_data(physeq.beta.100),"data.frame"))
```

```{r}
pslayer = plot_ordination(physeq.beta.100, physeq_clr_ord, type = "samples", color = "empo_2", shape = "empo_1") + ggtitle("Streptomyces Beta Diversity (PCoA) by EMPO level2") + 
  annotate("text", x = 0.38, y = 0.54, label = "PERMANOVA, p = 0.001") + 
  geom_point(size=1) + scale_colour_manual(values = c("#E69F00", "grey", "#009E73","#56B4E9")) + 
  theme_pubr(legend = "right", border = TRUE) + 
  theme(plot.title = element_text(hjust = 0.5)) #+ stat_ellipse(geom = "polygon", type = "t", alpha = 1/12, aes(fill = Layer)) 

pslayer 
```

```{r}
pdf("./Figures/Beta_div/Strep_BetaDiv_100.pdf", width = 8, height = 5 )
pslayer
dev.off()
```

